Alignment of Light and Mass in Lensing Galaxies

### 2014-05-05 shapes/paper.tex
Added bullets what needs still needs to be done and upload of a now mostly rather old version I worked on 1.5 years ago

% Open Tasks
%
% --- General ---
% - Add keywords
% - Check universities' names
% - Correct typos and reformulate some paragraphs
% - ...
%
%
% --- Introduction ---
% - Citiations needed in parametric vs. nonparametric comparison part
% - Citiations needed on effect of environment
% - Section on how this work might be able to rule out certain alternative gravity models
% - Describe theoretical expectations (ellipticities of dm halo and luminous part, misalignments, effect of environment)
% - ...
%
%
% --- Data ---
% - More detailed description how the baryionic mass distribution was measured
% - More details needed on galaxies in the sample? I don't think so though
% - ...
%
%
% --- Method ---
% - Maybe more details on GLASS and how it works (cite Jonathan's upcoming paper)
% - Plot showing how well the shape informations is recovered when applied to fake data
% - Citation on why not monotonically decreasing radial density profiles are instable
% - Generally look through method part
% - ...
%
%
% --- Results ---
% - Interpretation of 
% - Interpretation of ellipticities plot (why it is relevant that dm halos are significantly rounder)
% - Interpretation of misalignment versus ellipticity plot ('forbidden region', alignment with semi-minor axis, very elliptical galaxies well aligned)

@@

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Department of Biology and Health, Versailles Saint-Quentin-en-Yvelines University, France
}

@@

\textbf{Context/Background/Motivation}
\begin{itemize}
\item Why interesting (lensing can constrain mass distribution (enclosed mass constraint independent from model with only few percent uncertainty e.g. Kochanek 1991?, Wambsganss \& Paczyński 1994?), insights into structure formation)
\item What is to be expected from theory (dm, alternative gravity)
\end{itemize}

\textbf{Anything done before?}
\begin{itemize}
\item Surely cite: \cite{1997ApJ...482..604K,1998ApJ...509..561K,2006ApJ...649..599K,2009ApJ...690..670T,2012ApJ...761..170G}
\end{itemize}

\textbf{New here}
\begin{itemize}
\item Statistics
\item Free-form modelling
\item group versus stellar misalignment with dm halo
\end{itemize}

@@

\section{Data}\label{sec:data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. \citep{2011ApJ...740...97L})
\item Write why specifically these lenses were chosen, where they're from
\end{itemize}

@@

\section{Method}\label{sec:method}
\textbf{GLASS}
\begin{itemize}
\item Explain GLASS briefly and link to fake data tests
\end{itemize}

\subsection{Priors and assumptions}
\begin{itemize}
\item Explain what priors were chosen to model the lenses (Basically always pixrad 12, shear left free for quads, shear limited for doubles to 0.01, local gradient 50 degrees, smoothening prior 2, time delays if known. Exceptions are few. For some quads symmetry prior, for one double 0.01 was not possible therefore chose 0.1.)
\end{itemize}

\textbf{Shape measure}
\begin{itemize}
\item Explain shape measure
\item Link to tests on fake data
\item Include one plot of fake data test -> Jonathan
\item Illustrate everything with example lens PG1115.
\end{itemize}

@@

\textbf{Priors and assumptions}
We use different priors and assumptions to eliminate unphysical mass reconstructions of the lens. First of all, the densities $\kappa_{n}$ need to be positive. As mass distributions with positive radial gradients are instable (CITATION NEEDED!), the density must always increase towards the center. Therefore, we require the local density gradient in every pixel except the central one to point towards the center within an angle of $50^{\circ}$. The radial density profile decreases monotonically, as required.

@@

\textbf{Shape measure}

@@

Figure \ref{fig:1115rec} shows example plots of the reconstruction of the lens \textit{PG1115} illustrating this shape measure. In this lens system, there is a significant misalignment between the two mass distributions. 1$\sigma$- and 2$\sigma$-uncertainties on the angle estimates are shown. These errors are purely statistical, estimated by the ensemble of models. Figure \ref{fig:1115rec} also shows the arrival time surface and total, dark, and baryonic mass distributions.

\section{Results}\label{sec:results}
\begin{itemize}
\item Show some examples (stellar and dm contour plots, image of lens, arrival time surface, included mass)
\item Show money plots
\item Comment on a few lenses special lenses
\item Environmental effect?
\end{itemize}

@@

\begin{figure*}
 \begin{center}%\hspace{-7mm}\hspace{0.8\textwidth}
  \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_a.eps}
  \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_b.eps}
  \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_c.eps}\\
  \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_d.eps}
  \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_h.eps}\\
  \caption{Same as Figure \ref{fig:1115rec} but using time delay estimates from \cite{1997ApJ...489...21B}.}
  \label{fig:barkana}
 \end{center}
\end{figure*}

@@

\textbf{Misalignments between stellar and dark matter distributions}
Figure \ref{fig:moneyplots} shows the differences between the angles of the stellar and the dark matter semi-major axes versus the ratio of the included dark matter and total mass. For lenses with a significant stellar component inside the outermost image radius, the distributions seem to be rather aligned. It seems that if the luminous mass is comparable to the dark matter mass, the distributions align. For a smaller stellar contribution this is not necessarily the case. The scatter increases and there can be significant misalignments, which can be seen especially clearly in the systems \textit{PG1115} (also Figure \ref{fig:1115rec}) and \textit{RXJ0911}. While alignments between the distributions are preferred, they do not seem to be the rule.

@@

\textbf{Ellipticities of stellar and dark matter distributions}
The ellipticities of the distributions versus the mass ratio are shown in Figure \ref{fig:ellipticities}. We see that every dark matter halo has a lower ellipticity $\lambda_{1}/\lambda_{2}$ and is thus closer to sphericity than the luminous part of the galaxy. The systems \textit{B0712}, \textit{B1152} \textit{HE1104}, and \textit{MG2016} show especially large differences. It is not clear what could cause these discrepancies.

@@

\textbf{Misalignments between the dark matter distributions and environment}
Figure \ref{fig:environment} we look at the alignment of the semi-major axes of the dark matter halos and vectors pointing to the centroid of the group or cluster or a companion galaxy. Although the error bars are huge, we see that five of these ten lenses, \textit{Q0047}, \textit{B0712}, \textit{Q0957}, \textit{PG1115}, and \textit{B1422}, are probably aligned or at least consistent with the environment. The other half show slight misalignment.

@@

\section{Conclusion}\label{sec:conclusion}
\begin{itemize}
\item Recapitulate what were the main results
\item Answer whether this method is promising
\item What is needed to understand such misalignments better (-> Surely need to understand environment effects better, ...)
\end{itemize}

@@

>>>>

@@

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{1,2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Department of Biology and Health, Versailles Saint-Quentin-en-Yvelines University, France
}

@@

% \textbf{Context/Background/Motivation}
% \begin{itemize}
% \item Why interesting (lensing can constrain mass distribution (enclosed mass constraint independent from model with only few percent uncertainty e.g. Kochanek 1991?, Wambsganss \& Paczyński 1994?), insights into structure formation)
% \item What is to be expected from theory (dm, alternative gravity)
% \end{itemize}

% \textbf{Anything done before?}
% \begin{itemize}
% \item Surely cite: \cite{1997ApJ...482..604K,1998ApJ...509..561K,2006ApJ...649..599K,2009ApJ...690..670T,2012ApJ...761..170G}
% \end{itemize}

% \textbf{New here}
% \begin{itemize}
% \item Statistics
% \item Free-form modelling
% \item group versus stellar misalignment with dm halo
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Understand galaxy structure ($\rightarrow$ Links to galaxy evolution?)
\item Relevant for weak lensing (Intrinsic alignments)
\item Strong lensing reacts purely due to total mass distribution $\rightarrow$ Can disentangle light and mass
\item Free-form modelling (or \#constraints $\ll$ \#dof)
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rewrite introduction and change pitch to: Understand mass and light alignment, understand effect on IA
\end{itemize}

@@

\section{Data}\label{sec:data}
% \begin{itemize}
% \item SPS (Ignacio Ferreras, Dominik, etc. \citep{2011ApJ...740...97L})
% \item Write why specifically these lenses were chosen, where they're from
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Describe data set (why this data set, special features of galaxies (environment: y/n/unknonwn, elliptical/disk))
\item Maybe lose a couple of words on SPS?
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rephrase
\item Add maybe something on SPS
\end{itemize}

@@

\section{Method}\label{sec:method}
\subsection{GLASS}
% \textbf{GLASS}
% \begin{itemize}
% \item Explain GLASS briefly and link to fake data tests
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Describe code briefly
\item Mention PixeLens and two other GLASS papers
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rephrase
\item To avoid repetition, check what is really relevant for this paper, and for what we can just reference the other two GLASS papers
\item Check that same parameter names are used as in the other papers
\end{itemize}

@@

\subsection{Priors and assumptions}
% \subsection{Priors and assumptions}
% \begin{itemize}
% \item Explain what priors were chosen to model the lenses (Basically always pixrad 12, shear left free for quads, shear limited for doubles to 0.01, local gradient 50 degrees, smoothening prior 2, time delays if known. Exceptions are few. For some quads symmetry prior, for one double 0.01 was not possible therefore chose 0.1.)
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Describe what priors and assumptions we used
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rephrase
\item Maybe merge this subsection with previous subsection
\end{itemize}

We use different priors and assumptions to eliminate unphysical mass reconstructions of the lens. First of all, the densities $\kappa_{n}$ need to be positive. As mass distributions with positive radial gradients are instable (CITATION NEEDED!), the density must always increase towards the center. Therefore, we require the local density gradient in every pixel except the central one to point towards the center within an angle of $50^{\circ}$. The radial density profile decreases monotonically, as required.

@@

\subsection{Shape measure}
% \textbf{Shape measure}
% \begin{itemize}
% \item Explain shape measure
% \item Link to tests on fake data
% \item Include one plot of fake data test -> Jonathan
% \item Illustrate everything with example lens PG1115.
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Describe shape measure we use (same one as second GLASS paper)
\item Say that it has been tested with mock lenses (second GLASS paper)
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rephrase
\item Add formula how shape is defined
\item Summarize results of test of second GLASS paper
\end{itemize}

@@

% Figure \ref{fig:1115rec} shows example plots of the reconstruction of the lens \textit{PG1115} illustrating this shape measure. In this lens system, there is a significant misalignment between the two mass distributions. 1$\sigma$- and 2$\sigma$-uncertainties on the angle estimates are shown. These errors are purely statistical, estimated by the ensemble of models. Figure \ref{fig:1115rec} also shows the arrival time surface and total, dark, and baryonic mass distributions.

\section{Results}\label{sec:results}
% \begin{itemize}
% \item Show some examples (stellar and dm contour plots, image of lens, arrival time surface, included mass)
% \item Show money plots
% \item Comment on a few lenses special lenses
% \item Environmental effect?
% \end{itemize}
\textbf{Content:}
\begin{itemize}
\item Illustrate shape measure and reconstruction on PG1115 (with Tsevtkova time delays)
\item Describe different misalignment plots
\item Speculate what the plots could mean
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Start from scratch for this section
\item Use suggested structure (or think about better one)
\item Selection of plots is probably fine, but both data points for the two different time delay estimates of PG1115
\end{itemize}

\textbf{Structure of this section}
\begin{itemize}
\item Reconstruction of PG1115
\item Different ellipticities dark matter and stellar distribution
\item Misalignment stellar and dark matter distributions
\item Alignment dark matter and group centroid (maybe merge with previous subsection)
\end{itemize}

@@

% \begin{figure*}
%  \begin{center}%\hspace{-7mm}\hspace{0.8\textwidth}
%   \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_a.eps}
%   \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_b.eps}
%   \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_c.eps}\\
%   \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_d.eps}
%   \includegraphics[height=0.32\textwidth]{Figures/1115_barkana_h.eps}\\
%   \caption{Same as Figure \ref{fig:1115rec} but using time delay estimates from \cite{1997ApJ...489...21B}.}
%   \label{fig:barkana}
%  \end{center}
% \end{figure*}

@@

\subsection{Misalignments between stellar and dark matter distributions}
Figure \ref{fig:moneyplots} shows the differences between the angles of the stellar and the dark matter semi-major axes versus the ratio of the included dark matter and total mass. For lenses with a significant stellar component inside the outermost image radius, the distributions seem to be rather aligned. It seems that if the luminous mass is comparable to the dark matter mass, the distributions align. For a smaller stellar contribution this is not necessarily the case. The scatter increases and there can be significant misalignments, which can be seen especially clearly in the systems \textit{PG1115} (also Figure \ref{fig:1115rec}) and \textit{RXJ0911}. While alignments between the distributions are preferred, they do not seem to be the rule.

@@

\subsection{Ellipticities of stellar and dark matter distributions}
The ellipticities of the distributions versus the mass ratio are shown in Figure \ref{fig:ellipticities}. We see that every dark matter halo has a lower ellipticity $\lambda_{1}/\lambda_{2}$ and is thus closer to sphericity than the luminous part of the galaxy. The systems \textit{B0712}, \textit{B1152} \textit{HE1104}, and \textit{MG2016} show especially large differences. It is not clear what could cause these discrepancies.

@@

\subsection{Misalignments between the dark matter distributions and environment}
Figure \ref{fig:environment} we look at the alignment of the semi-major axes of the dark matter halos and vectors pointing to the centroid of the group or cluster or a companion galaxy. Although the error bars are huge, we see that five of these ten lenses, \textit{Q0047}, \textit{B0712}, \textit{Q0957}, \textit{PG1115}, and \textit{B1422}, are probably aligned or at least consistent with the environment. The other half show slight misalignment.

@@

\section{Conclusion}\label{sec:conclusion}
\textbf{Content:}
\begin{itemize}
\item Quick overview of what was presented
\item State again main results
\item Say why this results are interesting
\item What is needed to understand such misalignments better (-> Surely need to understand environment effects better, ...)
\end{itemize}

\textbf{To Do:}
\begin{itemize}
\item Rephrase
\item Maybe wait for final results
\end{itemize}

@@

%%%%%%%% IGNORE THIS %%%%%%%%%%
% Open Tasks
%
% --- General ---
% - Add keywords
% - Correct typos and reformulate some paragraphs
% - ...
%
%
% --- Introduction ---
% - Citiations needed in parametric vs. nonparametric comparison part
% - Citiations needed on effect of environment
% - Section on how this work might be able to rule out certain alternative gravity models
% - Describe theoretical expectations (ellipticities of dm halo and luminous part, misalignments, effect of environment)
% - ...
%
%
% --- Data ---
% - More detailed description how the baryionic mass distribution was measured
% - More details needed on galaxies in the sample? I don't think so though
% - ...
%
%
% --- Method ---
% - Maybe more details on GLASS and how it works (cite Jonathan's upcoming paper)
% - Plot showing how well the shape informations is recovered when applied to fake data
% - Citation on why not monotonically decreasing radial density profiles are instable
% - Generally look through method part
% - ...
%
%
% --- Results ---
% - Interpretation of 
% - Interpretation of ellipticities plot (why it is relevant that dm halos are significantly rounder)
% - Interpretation of misalignment versus ellipticity plot ('forbidden region', alignment with semi-minor axis, very elliptical galaxies well aligned)
